\section*{Old Introduction}

This work revisits {\sl set reconciliation}, a problem consisting in
synchronizing two multisets while minimizing the amount of data transmitted. Set
reconciliation arises in many practical situations, the most typical of which is
certainly incremental backups performed over a slow network link.\smallskip

Several efficient and elegant solutions are known to achieve set reconciliation of multisets containing atomic elements of a fixed size. For instance, \cite{PSRec} manages to perform set reconciliation using a bandwidth which is linear in the size of the symmetric difference of the multisets multiplied by the size of the elements, which is optimal in this setting. We refer the reader to~\cite{PSRec,Mins1,Whats} for more on this problem's history and its existing solutions.\smallskip

However, in the case where the elements to be synchronized can be very large (e.g., files during a backup), we must use checksums to identify the differing files before transferring them, and the question of the size of the checksum to use is non-trivial. In this article, we propose a new reconciliation protocol called ``Divide \& Factor'' ({\sc d\&f}) based on number theory. In terms of asymptotic transmission complexity, the proposed procedure reaches optimality as well. 
\comm{Fabrice@Antoine: Je ne suis pas tout à fait d'accord avec ce paragraphe, ton abstract est beaucoup mieux...}
In addition, the new protocols offer a very interesting gamut of parameter trade-offs. We provide an analysis of the protocol's complexity in terms of transmission and computation, as well as a probabilistic analysis of the possible choices of checksum sizes; we also provide an implementation of the protocol and experimental results.\smallskip

This paper is structured as follows: Section~\ref{dandf} presents ``Divide \&
Factor'', our set reconciliation protocol. Section~\ref{trans} presents the
transmission complexity of the protocol, introduces two transmission
optimizations and analyzes them in detail. Section \ref{comp} analyzes the
computational complexities of the proposed protocols. Section~\ref{files}
explain how to use a set reconciliation algorithm to perform file
synchronization. Section~\ref{program} presents our implementation \btrsync and
reports practical experiments and benchmarks against the popular software
\rsync.\smallskip


\subsection{USELESS --- How to Stop a Probabilistic Washing Machine?} 
\comm{Remplacer $\ell$ par $\lambda$}
We now combine both optimizations and assume that $\ell$ laundry rounds are necessary for completing some given reconciliation task using a half-sized $p$. By opposition to Section~\ref{sec:basic}, confirming correct protocol termination is now non-trivial.
\comm{Fabrice: oups, pas sûr de tout comprendre ici, il faudrait que l'on en rediscute... En particulier, le CRT n'a pas d'erreur a priori, sauf erreur de ma part...}

\comm{La définition de $\zeta(u)$ est assez gratuite vu que ça n'intervient pas dans la suite.}
We say that a \textit{round failure} occurs whenever a round results in an $(a',b')\neq (a,b)$ satisfying Equation~\eqref{eqab}. Let the round failure probability be some function $\zeta(u)$ (that we did not estimate). If $u$ is kept small (for efficiency reasons), the probability $\left(1-\zeta(u)\right)^{\ell}$ that the protocol will properly terminate may dangerously drift away from one.

If $v$ of $\ell+v$ rounds fail, Oscar needs to solve a problem called \textit{Chinese Remaindering With Errors}~\cite{phong}:

\begin{problem}{\it (Chinese Remaindering With Errors Problem: {\sc crwep}).} Given as input integers $v$, $B$ and $\ell+v$ points $(s_1,p_1),\ldots,(s_{\ell+v},p_{\ell+v})\in \mathbb{N}^2$ where the $p_i$'s are coprime, output all numbers $0 \leq s < B$ such that $s \equiv s_i \bmod p_i$ for at least $\ell$ values of $i$.
\end{problem}

We refer the reader to~\cite{phong} for more on this problem, which is beyond our scope. Boneh~\cite{boneh} provides a polynomial-time algorithm for solving the {\sc crwep} under certain conditions satisfied in our setting.

But how can we confirm the solution? As mentioned in section \ref{reco}, Neil will send to Oscar $H=\Hash(\mathfrak{F}')$ as the interaction starts. As long as Oscar's {\sc crwep} resolution will not yield a state matching $H$, the parties will continue the interaction.



\section{Old D\&F Set Reconciliation Section}

Section \ref{basic} presents a basic
version of the proposed protocol. This basic version suffers from two
limitations: it works only if the number of differences to reconcile is bound
and it may fail leave the synchronized party in an erroneous state. Failure
avoidance is overcome in section \ref{reco} and an extension to arbitrary
numbers of differences is given in section \ref{insuf}.

\subsection{Problem Definition and Notations}

\underline{O}scar possesses an \underline{o}ld version of a directory $\mathfrak{D}$ that he wishes to update. \underline{N}eil has the \underline{n}ew, up-to-date version $\mathfrak{D}'$: $\mathfrak{D}$ and
$\mathfrak{D}'$ can differ both in their files and in their tree structures. Oscar wishes to obtain $\mathfrak{D}'$ but {\sl exchange as little data as possible} during the synchronization process.\smallskip

To tackle this problem we separate the {\sl ``what''} from the {\sl ``where''} by considering files as a tuple of their location and content. 
In other words, we will first synchronize all the file contents and then move files to the adequate location. 
\comm{Fabrice: NON, ce n'est pas vrai: on considère vraiment qu'un fichier est ``path + content''}
We consider that $\mathfrak{D}$ is a multiset of files which we denote as $\mathfrak{F}=\{F_0,\ldots,F_{n}\}$, and likewise represent $\mathfrak{D'}$ as $\mathfrak{F}'=\{F'_0,\ldots,F'_{n'}\}$.\smallskip
\comm{Fabrice: Qu'est-ce que $\mathfrak{D}$ ? Est-ce $\mathfrak{F}$ ? Dans tous les cas, il s'agit d'un set, vu la représentation indiquée}

Let $T$ be the number of discrepancies between $\mathfrak{F}$ and $\mathfrak{F}'$ that Oscar wishes to learn, {\sl i.e.} the symmetric difference of $\mathfrak{F}$ and $\mathfrak{F}'$:

$$T=\#\mathfrak{F}+\#\mathfrak{F}'-2 \#\left(\mathfrak{F} \bigcap \mathfrak{F}'\right)=\#\left(\mathfrak{F}\bigcup\mathfrak{F}'\right)-\#\left(\mathfrak{F}\bigcap\mathfrak{F}'\right)$$

Given a file $F$, we denote by $\mbox{{\tt Hash}}(F)$ its image by a collision-resistant hash function such as {\sc sha}-1.\comm{Fabrice: ${\tt Hash}$ can be introduced latter when needed (section 2.3)} Let $\mbox{{\tt HashPrime}}(F)$\footnote{The design of \mbox{{\tt HashPrime}} is addressed in Appendix \ref{sec:hashprime}.} be a function hashing files (uniformly) into primes smaller than $2^u$ for some $u\in \mathbb{N}$. Define the shorthand notations: $h_i=\mbox{{\tt HashPrime}}(F_i)$ and $h'_i=\mbox{{\tt HashPrime}}(F'_i)$.\smallskip

\subsection{Description of the Basic Exchanges}
\label{basic}

The number of differences $T$ is unknown to Oscar and Neil. However, for the time being, we will assume that $T$ is smaller than some $t$ and attempt to perform
synchronization. If $T \leq t$, synchronization will succeed; if $T > t$ the parties will transmit more information later to complete the synchronization, as explained in section \ref{insuf}.

We generate a prime $p$ such that:

\begin{equation}
\label{equp}
2^{2ut} \leq p < 2^{2ut+1}
\end{equation}

TODO: such a prime exists (reference Chebychev, 1840).

Given $\mathfrak{F}$, Oscar generates and sends to Neil the redundancy:

$$
c=\prod_{F_i\in \mathfrak{F}} \mbox{{\tt HashPrime}}(F_i)=\prod_{i=1}^n h_i \bmod p
$$

Neil computes:\smallskip

$$c'=\prod_{F'_i\in \mathfrak{F'}} \mbox{{\tt HashPrime}}(F'_i)=\prod_{i=1}^{n'} h'_i \bmod p{~~~\mbox{and}~~~}s=\frac{c'}{c} \bmod p$$

Using~\cite{vallee}\comm{Fabrice: Quel est l'intérêt de cette citation ???} the integer $s$ can be written as:
$$s=\frac{a}{b} \bmod p{\mbox{~where the~}G_i\mbox{~denote files and~}}
\left\{
\begin{array}{lcr}
a & =&  \prod\limits_{G_i \in \mathfrak{F}'\wedge G_i \not\in\mathfrak{F}} \mbox{{\tt HashPrime}}(G_i) \\
\\
b & = & \prod\limits_{G_i \not\in\mathfrak{F}' \wedge G_i \in\mathfrak{F}} \mbox{{\tt HashPrime}}(G_i)
\end{array}
\right.
$$
\comm{Fabrice: je ne comprends pas le ``where the $G_i$ denote files...''}
Note that if our assumption $T \leq t$ is correct, $\mathfrak{F}$ and $\mathfrak{F}'$ differ by at most $t$ elements and $a$ and $b$ are strictly less than $2^{ut}$. The problem of recovering $a$ and $b$ from $s$ efficiently is known as {\sl Rational Number Reconstruction}~\cite{pan2004rational,wang2003acceleration}.
 theorem \ref{theo} (see~\cite{cryptorational}) guarantees that it can be solved in this setting.
The following theorem is a slightly modified version of Theorem~1 in \cite{cryptorational}:
\begin{theorem}
\label{theo}
Let $a,b \in {\mathbb Z}$ two co-prime integers such that $0 \leq a \leq A$ and $0<b \leq B$. Let $p>2AB$ be a prime and $s=a b^{-1} \bmod p$. Then $a,b$ are uniquely defined given $s$ and $p$, and can be recovered from $A,B,s,p$ in polynomial time.
\end{theorem}

Taking $A=B=2^{ut}-1$, Equation \eqref{equp} implies that $AB<p$. Moreover, $0 \leq a \leq A$ and $0 <b \leq B$. Thus Oscar can recover $a$ and $b$ from $s$ in polynomial time: a possible option is to use Gauss algorithm for finding the shortest vector in a bi-dimensional lattice~\cite{vallee}.
\comm{Fabrice: certes, mais on utilise directement un Euclide étendu tronqué. Et pourquoi citer Vallée qui est un peu incompréhensible dans notre cas... TODO: check that in our program we ensure $a$ and $b$ co-prime !! otherwise, it may fail !!!!!}
By testing the divisibility of $a$ and $b$ by the $h_i$ and the $h'_i$, Neil and Oscar can attempt to identify the discrepancies between $\mathfrak{F}$ and $\mathfrak{F}'$ and settle them.\smallskip

\begin{figure}
\begin{center}
\begin{tabular}{|lcl|}\hline
~~{\bf Oscar}                       &                                                      &   {\bf Neil}~\\
~~compute $c$&                                                      &\\
                                   &~~{{\LARGE $\stackrel{c}{\longrightarrow}$}}~~        &   \\
                                   &                                                      &compute $a,b$~\\
                                   &                                                      &if $a$ doesn't factor as a product of $h'_i$s~~\\
                                   &                                                      &~~~~then output $\bot_{\mbox{{\tiny {\sf bandwidth}}},1}$ and halt~~\\
                                   &                                                      &$\mathfrak{S}\leftarrow\{F'_i \mbox{~s.t.~} a \bmod h'_i =0\}$~~\\
                                   &~~{\LARGE $\stackrel{\mathfrak{S},b}{\longleftarrow}$}&\\
~~if $b$ doesn't factor as a product of $h_i$'s&&\\
~~~~~~then output $\bot_{\mbox{{\tiny {\sf bandwidth}}},2}$ and halt &&\\
~~delete files s.t. $b \bmod h_i =0$&                                                      &\\
~~add $\mathfrak{S}$ to the disk    &                                                      &\\\hline
\end{tabular}
\end{center}
\caption{Basic Protocol.}\label{fig:one}
\end{figure}

The formal description of the protocol is given in Figure \ref{fig:one}. The ``output $\bot_{\mbox{{\tiny {\sf bandwidth}}},\square}$ '' protocol interruptions will:

\begin{itemize}
\item never occur if the assumption $T \leq t$ holds.

\item occur with high probability if $T > t$. Indeed, for a potential $\bot_{\mbox{{\tiny {\sf bandwidth}}},1}$ to be overlooked, the $ut$-bit number $a$ must perfectly factor over a set of $n$ primes of size $u$. If we assume that $a$ is ``random'', the probability $\gamma$ that $a$ is divisible by some $h_i$ is essentially $\gamma \sim 1/h_i \sim 2^{-u}$, the probability that $a$ is divisible by exactly $t$ digests is:
$$\alpha=\binom{n}{t} \gamma^t (1 - \gamma)^{n - t} \sim \binom{n}{t} 2^{-u t} (1 - 2^{-u})^{n - t}$$ and the probability that the protocol does not terminate by a $\bot_{\mbox{{\tiny {\sf bandwidth}}},\square}$ when $T > t$ is $\sim\alpha^2$.
\comm{However, we are not interested in the fact $a$ is divisible by ``exactly'' $t$ digests but the fact that $a$ can be factorized over of the basis of $h_i$...}
\end{itemize}

The very existence of $\bot_{\mbox{{\tiny {\sf bandwidth}}},\square}$'s is annoying for two reasons:
\begin{itemize}
\item A file synchronization procedure that works {\sl only} for a limited number of differences is not really useful in practice. Thus, section \ref{insuf} explains how to extend the protocol to perform the synchronization even when the number of differences $T$ exceeds the initial estimation $t$.\smallskip
\item If, by sheer bad luck, both $\bot_{\mbox{{\tiny {\sf
  bandwidth}}},\square}$'s went undetected (double accidental factorization) the
  Basic Protocol (Fig. \ref{fig:one}) may leave Oscar in an inconsistent state.
\end{itemize}

Double accidental factorization is not only possible source of inconsistent states: as we did not specifically require $\mbox{{\tt HashPrime}}$ to be collision-resistant, the events

$$
\begin{array}{lll}
{
\bot_{\mbox{{\tiny {\sf collision}}},1}=\left\{
\begin{array}{l}
h'_i = h'_j \mbox{~for~}i\neq j\\
\\
a \bmod h_i =0
\end{array}
\right.
}&\mbox{~~~and/or~~~}&{
\bot_{\mbox{{\tiny {\sf collision}}},2}=\left\{
\begin{array}{l}
h_i = h_j \mbox{~for~}i\neq j\\
\\
b \bmod h'_i =0
\end{array}
\right.}\\
\end{array}
$$

will cause Neil to send wrong files in $\mathfrak{S}$ ($\bot_{\mbox{{\tiny {\sf collision}}},1}$) and/or have Oscar unduely delete files owned by Neil ($\bot_{\mbox{{\tiny {\sf collision}}},2}$).\smallskip

Inconsistent states may hence stem from three events:

\begin{center}
\begin{tabular}{llll}
$\bullet$~~&\multicolumn{3}{l}{accidental double factorization of $a$ and/or $b$ when $T > t$ (probability $\alpha^2$)}\\
$\bullet$~~&$\bot_{\mbox{{\tiny {\sf collision}}},1}$   &$=$& collisions within the set $\{h'_i\}$\\
$\bullet$~~&$\bot_{\mbox{{\tiny {\sf collision}}},2}$ &$=$& collisions within the set $\{h_i\}$\\
\end{tabular}\smallskip
\end{center}
\comm{Fabrice: ce qui est un peu tordu, c'est que les collisions qui nous intéressent sont les collisions entre $\mathfrak{S}$ et $\mathfrak{F} \cup \mathfrak{F}'$... il faut qu'on en discute}

Section \ref{reco} explains how protect the protocol from all inconsistent events at once.

\subsection{Avoiding Inconsistency}
\label{reco}

\comm{Dire exactement ce que c'est qu'on avoid, et dire quand même qu'on a un risque minime de collision sur $H$ (mais que c'est juste un coût constant et on peut le prendre suffisamment grand pour que ce soit négligeable).}

The Basic Protocol of Figure \ref{fig:one} is fully deterministic. Hence if any sort of trouble occurs, repeating the protocol will be of no help. We modify the protocol as follows:

\begin{itemize}
\item Let $H\leftarrow\mbox{{\tt Hash}}(\mathfrak{F}')$\comm{Fabrice: This notation (hash of a set) is not defined... and it may be useful to recall the definition of ${\tt Hash}$ here (collision-resistant) if the reader has forgotten it...}
\item Replace $\mbox{{\tt HashPrime}}(F)$ by a diversified $\hbar_k(F)=\mbox{{\tt HashPrime}}(k|F)$.\smallskip
\item Define the shorthand notations: $\hbar_{k,i}=\hbar_k(F_i)$ and $\hbar'_{k,i}=\hbar_k(F'_i)$.\smallskip
\item Let $\mbox{{\sf StepProtocol}}(k)$ denote the sub-protocol shown in Figure \ref{fig:step}.
\item Use the protocol of Figure \ref{fig:itera} as a fully functional reconciliation protocol for $T \leq t$.
\end{itemize}

\begin{figure}
\begin{center}
\begin{tabular}{|lcl|}\hline
~~{\bf Oscar}                       &                                                      & {\bf Neil}~\\
                                   &                                                       &if $a$ doesn't factor as a product of $\hbar'_{k,i}$s~~\\
                                   &                                                       &~~~~then output $\bot_{\mbox{{\tiny {\sf bandwidth}}},1}$ and halt~~\\
                                   &                                                       &$\mathfrak{S}\leftarrow\{F'_i \mbox{~s.t.~} a \bmod \hbar'_{k,i} =0\}$~~\\
  ~~                                 &                                                       &if there are collisions between $\mathfrak{S}$ and $\mathfrak{F}$~~\\
                                   &                                                       &~~~~then output $\bot_{\mbox{{\tiny {\sf collision}}},1}$ and halt~~\\
                                   &~~{\LARGE $\stackrel{\mathfrak{S},b}{\longleftarrow}$} &\\
~~if $b$ doesn't factor as a product of $\hbar_{k,i}$'s&&\\
~~~~~~then output $\bot_{\mbox{{\tiny {\sf bandwidth}}},2}$ and halt &&\\
~~$\mathfrak{A}\leftarrow\{F_i \mbox{~s.t.~} b \bmod \hbar_{k,i} =0\}$ &&\\
~~if there are collisions in $\mathfrak{A}$ &                                   & \\
~~~~~~then output $\bot_{\mbox{{\tiny {\sf collision}}},2}$ and halt~~&                      &~~\\
~~if $H \neq \mbox{{\tt Hash}}(\mathfrak{F}\bigcup\mathfrak{S} - \mathfrak{A})$ &                                                      &\\
~~~~~~then output $\bot_{\mbox{{\tiny {\sf bandwidth}}},3}$ and halt &&\\
~~add $\mathfrak{S}$ to the disk and erase $\mathfrak{A}$ from the disk &                                                      &\\
~~return {\sf success} &                                                      &\\\hline
\end{tabular}
\end{center}
\caption{$\mbox{{\sf StepProtocol}}(k)$.}\label{fig:step}
\end{figure}

\subsubsection{Note:} To avoid transmitting the (potentially very voluminous) $\mathfrak{S}$ during {\sf StepProtocol} before knowing if one of the errors $\bot_{\mbox{{\tiny {\sf bandwidth}}},2},\bot_{\mbox{{\tiny {\sf bandwidth}}},3},\bot_{\mbox{{\tiny {\sf collision}}},2}$ will occur, Neil may transmit $$\mathfrak{S}'=\{\mbox{{\tt Hash}}(F'_i),~F'_i\in \mathfrak{S}\}$$ instead of $\mathfrak{S}$ and send $\mathfrak{S}$ only after successfully passing the $\bot_{\mbox{{\tiny {\sf bandwidth}}},3}$ test. The definition of $H$ must be changed accordingly to 

$$H=\mbox{{\tt Hash}}(\{\mbox{{\tt Hash}}(F'_i),~F'_i\in \mathfrak{F}'\})$$
\comm{Fabrice: en même temps, le Hash d'un ensemble, que l'on n'a pas définit, a tout intérêt à déjà être de cette forme, sinon, on a des problèmes pour avoir une concaténation ``propre''}

\begin{figure}
\begin{center}
\begin{tabular}{|lcl|}\hline
~~{\bf Oscar}                      &                                                      &   {\bf Neil}~\\
~~                                 &                                                      & compute $H\leftarrow\mbox{{\tt Hash}}(\mathfrak{F}')$\\
                                   &~~{{\LARGE $\stackrel{H}{\longleftarrow}$}}~~   &   \\
~~compute $c$&                                                                             &\\
                                   &~~{{\LARGE $\stackrel{c}{\longrightarrow}$}}~~         &   \\
                                   &                                                       &compute $a,b$~\\
                                   &                                                      &$k\leftarrow 1$~\\
                                   &~~{{\LARGE $\stackrel{\mbox{{\small{\sf StepProtocol}}}(k)}{\longleftarrow}$}}~~   &while $\mbox{{\sf StepProtocol}}(k)=\bot_{\mbox{{\tiny {\sf collision}}},\square}$~~\\
                                   &                                                      &~~~~$k\leftarrow k+1$~\\\hline
\end{tabular}
\end{center}
\caption{Fully Functional Protocol for $T \leq t$.}\label{fig:itera}
\end{figure}

\subsection{Handling a High Number of Differences}
\label{insuf}

To extend the protocol to an arbitrary $T$, Oscar and Neil agree on an infinite set of primes $p_1,p_2,\ldots$ As long as the protocol fails with a $\bot_{\mbox{{\tiny {\sf bandwidth}}},\square}$ status, Neil and Oscar redo the protocol with a new $p_\ell$ and Neil will keep accumulating information about the difference between $\mathfrak{F}$ and $\mathfrak{F}'$ as shown in Appendix \ref{sec:extended}. 
Each of this repetition is called a round.
Note that no information is lost and that the transmitted modular knowledge about the difference adds up until it reaches a threshold sufficient to reconcile $\mathfrak{F}$ and $\mathfrak{F}'$. \smallskip

More precisely, let us suppose $2^{2 u t_k} \le p_\ell < 2^{2 u t_k +1}$.
Let us write $P_k = p_1 \dots p_k$ and $T_k = u (t_1 + \dots t_k)$.
After receiving the redundancies $c_1,\dots,c_k$ corresponding to $p_1,\dots,p_k$, Neil has as many information as if Oscar had transmitted a redundancy $C_k$ corresponding to the modulo $P_k$, and can compute $S_k = C'_k / C_k$ from $s_k = c'_k/c_k$ and $S_{k-1}$ using the CRT (TODO ref ?).
Therefore, the number $\lambda$ of rounds used is the minimum number $k$ such that $T_k \ge T$.
If $t_1 = t_2 = \dots = t$, then $\lambda = \lceil t/T \rceil$.

All $\bot$ treatments were removed from Appendix \ref{sec:extended} for the sake of clarity (these can be very easily added by modifying Appendix \ref{sec:extended} {\sl mutatis mutandis}). In essence, the rules are: add information modulo a new $p_\ell$ whenever the protocol fails with a $\bot_{\mbox{{\tiny {\sf bandwidth}}},\square}$ and increment $k$ whenever the protocol fails with a $\bot_{\mbox{{\tiny {\sf collision}}},\square}$.\smallskip
\comm{Fabrice: plus clair si on le met dans une deuxième figure en appendix quand même, je pense...}

A typical execution sequence is thus expected to be something like:

$$\bot_{\mbox{{\tiny {\sf bandwidth}}},1},\bot_{\mbox{{\tiny {\sf bandwidth}}},1},\bot_{\mbox{{\tiny {\sf bandwidth}}},1},\bot_{\mbox{{\tiny {\sf bandwidth}}},1},\bot_{\mbox{{\tiny {\sf collision}}},1},\bot_{\mbox{{\tiny {\sf collision}}},1},\mbox{{\sf success}}$$

\section{Extended Protocol}
\label{sec:extended}

\begin{center}
\begin{tabular}{lcl}\toprule
\multicolumn{3}{c}{{\sf First phase during which Neil amasses modular
information on the difference~~}} \\\midrule
~~{\bf Oscar}                      &                                                      &   {\bf Neil}~\\
                                   &                                                      &start the protocol with $p_1$~\\
                                   &~~{{\LARGE $\stackrel{c_1}{\longrightarrow}$}}~~      &   \\
                                   &                                                      &computes $a,b$ using $p_1$~\\
                                   &                                                      &if $a$ factors properly then go to {\sf Final Phase}\\
                                   &                                                      &~~~~~~else perform the protocol with $p_2$~~\\
                                   &~~{{\LARGE $\stackrel{c_2}{\longrightarrow}$}}~~      &   \\
                                   &                                                      &computes $c \bmod p_1 p_2=\mbox{CRT}_{p_1,p_2}(c_1,c_2)$~~\\
                                   &                                                      &computes $a,b$ using $p_1 p_2$~\\
                                   &                                                      &if $a$ factors properly then go to {\sf Final Phase}\\
                                   &                                                      &~~~~~~else perform the protocol with $p_3$~~\\
                                   &~~{{\LARGE $\stackrel{c_3}{\longrightarrow}$}}~~      &   \\
                                   &                                                      &computes $c \bmod p_1 p_2 p_3=\mbox{CRT}_{p_1,p_2,p_3}(c_1,c_2,c_3)$~~\\
                                   &                                                      &computes $a,b$ using $p_1 p_2 p_3$~\\
                                   &                                                      &if $a$ factors properly then go to {\sf Final Phase}\\
                                   &                                                      &~~~~~~else perform the protocol with $p_4$ ~~\\
                                   &                  $\vdots$
& \\\midrule
\multicolumn{3}{c}{{\sf Final Phase~~}} \\\midrule
                                   &                                                      & \\
                                   &                                                      &Let $\mathfrak{S}=\{F'_i \mbox{~s.t.~} a \bmod h'_i =0\}$~~\\
                                   &~~{\LARGE $\stackrel{\mathfrak{S},b}{\longleftarrow}$}&\\
                                   ~~deletes files s.t. $b \bmod h_i =0$&                                                      &\\
                                   ~~adds $\mathfrak{S}$ to the disk
                                   &
                                   &\\\bottomrule
\end{tabular}
\end{center}

Note that parties do not need to store the $p_i$'s in full. Indeed, the $p_i$s could be subsequent primes sharing their most significant bits. This reduces storage per prime to a very small additive constant $ \cong \mbox{ln}(p_i) \cong \mbox{ln}(2^{2tu+2}) \cong 1.39(tu+1)$ of about $\log_2(tu)$ bits.

\section{Power of Two Protocol}
\label{powtwo}

In this variant Oscar computes $c$ in $\mathbb{N}$:

$$
c=\prod_{F_i\in \mathfrak{F}} \mbox{{\tt HashPrime}}(F_i) = \prod_{i=1}^n h_i \in \mathbb{N}
$$

and considers $c=\bar{c}_{n-1}|\ldots|\bar{c}_2|\bar{c}_0$ as the concatenation of $n$ successive $u$-bit strings. Again, we omit the treatment of $\bot$s for the sake of clarity.

\begin{center}
\begin{tabular}{lcl}\toprule
\multicolumn{3}{c}{{\sf First phase during which Neil amasses modular
information on the difference~~}} \\\midrule
~~{\bf Oscar}                      &                                                      &   {\bf Neil}~\\
~~computes $c\in \mathbb{N}$       &                                                      &\\
                                   &~~{{\LARGE $\stackrel{\bar{c}_0}{\longrightarrow}$}}~~&   \\
                                   &                                                      &computes $a,b$ modulo $2^u$~\\
                                   &                                                      &if $a$ factors properly then go to {\sf Final Phase}~~\\
                                   &                                                      &~~~~~~else request next chunk $\bar{c}_1$~~\\
                                   &~~{{\LARGE $\stackrel{\bar{c}_1}{\longrightarrow}$}}~~&   \\
                                   &                                                      &construct $c \bmod 2^{2u}=\bar{c}_1|\bar{c}_0$~~\\
                                   &                                                      &computes $a,b$ modulo $2^{2u}$~\\
                                   &                                                      &if $a$ factors properly then go to {\sf Final Phase}~~\\
                                   &                                                      &~~~~~~else request next chunk $\bar{c}_2$~~\\
                                   &~~{{\LARGE $\stackrel{\bar{c}_2}{\longrightarrow}$}}~~      &   \\
                                   &                                                      &construct $c\bmod 2^{3u}=\bar{c}_2|\bar{c}_1|\bar{c}_0$~~\\
                                   &                                                      &computes $a,b$ modulo $2^{3u}$~\\
                                   &                                                      &if $a$ factors properly then go to {\sf Final Phase}~~\\
                                   &                                                      &~~~~~~else request next chunk $\bar{c}_3$~~\\
                                   &                  $\vdots~(\mbox{~for~}2t
\mbox{~rounds~})$    & \\\midrule
\multicolumn{3}{c}{{\sf Final Phase~~}} \\\midrule
                                   &                                                      & \\
                                   &                                                      &Let $\mathfrak{S}=\{F'_i \mbox{~s.t.~} a \bmod 2^{2tu} =0\}$~~\\
                                   &~~{\LARGE $\stackrel{\mathfrak{S},b}{\longleftarrow}$}&\\
                                   ~~deletes files s.t. $b \bmod 2^{2t u} =0$&        &\\
                                   ~~adds $\mathfrak{S}$ to the disk
                                   &
                                   &\\\bottomrule
\end{tabular}
\end{center}




%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "btrsync"
%%% End: 
